Creating optimal conditions for reproducible data analysis in \R with `fertile'\protect\thanks{The authors gratefully acknowledge contributions from Hadley Wickham, Jenny Bryan, Greg Wilson, Edgar Ruiz, and other members of the \pkg{tidyverse} team.}

### 2020-08-05 STAT-paper-template/APA/fertile.Rnw
expanded results + other sections

@@

\begin{enumerate}
\item Identify an \R function that is likely to be involved in operations that may break reproducibility. Popular functions associated with only one package (e.g., \func{read\_csv} from \pkg{readr}) are ideal candidates.
\item Create a function in \pkg{fertile} with the same name that takes the same arguments (and always the dots \cmd{...}).
\item Write this new function so that it: 1) captures any arguments; 2) logs the name of the function called; 3) perform any checks on these arguments; and finally 4) calls the original function with the original arguments. Except where warranted, the execution looks the same to the user as if they were calling the original function.
\end{enumerate}
Most shims are quite simple and look something like what is shown below for \func{read\_csv}.

@@

In order to ensure that the \pkg{fertile} versions of functions ("shims") are always favored ("mask") over their original namesakes when called, \pkg{fertile} uses its own shims of the \func{library} and \func{require} functions to  manipulate the \R \func{search} path so that it is always located in the first position. In the \pkg{fertile} version of \func{library}, we detach \pkg{fertile} from the search path, load the requested package, and then re-attach \pkg{fertile}. This ensures that when a user executes a command, \R will check \pkg{fertile} for a matching function before considering other packages.

@@

<<is_file>>=
is_data_file(fs::path(proj_dir, "mice.csv"))
is_image_file(fs::path(proj_dir, "mice.csv"))
is_image_file(fs::path(proj_dir, "proteins_v_time.png"))
@

@@

<<sandbox>>=
dir <- getwd()
new_dir <- sandbox(dir)
dir
fs::dir_ls(dir) %>% head(3)
new_dir
fs::dir_ls(new_dir) %>% head(3)
@

@@

<<proj_pkg_script, eval=TRUE>>=
install_script <- proj_pkg_script(proj_dir)
cat(readChar(install_script, 1e5))
@

@@

\pkg{fertile} is designed to: 1) be simple enough that users with minimal \R experience can use the package without issue; 2) increase the reproducibility of work produced by its users, and; 3) educate its users on why their work is or is not reproducible and provide guidance on how to address any problems.

To test \pkg{fertile}'s effectiveness, we began an initial randomized control trial of the package on an introductory undergraduate data science course at Smith College in Spring 2020 \footnote{This study was approved by Smith College IRB, Protocol \#19-032}. Unfortunately, the trial had to be postponed after the COVID-19 pandemic significantly altered the experimental conditions at the midpoint of testing.
Although the experiment was unsuccessful in its first attempt, we hope to run the same trial again and gather data on \pkg{fertile}'s effectiveness.

@@

\item In order to test \pkg{fertile}'s effectiveness, both groups are given a short quiz on different components of reproducibility that are intended to be taught by \pkg{fertile} at both the beginning and end of the semester. Their scores are then compared to see whether one group learned more than the other group or whether their scores were essentially equivalent.

@@

>>>>

Each test can also be run individually by calling the function matching its check name:

<<individual checks, warning = FALSE>>=

has_tidy_raw_data(proj_dir)

@@

\begin{enumerate}
\item Identify an \R function that is likely to be involved in operations that may break reproducibility. Popular functions associated with only one package (e.g., \func{read\_csv} from \pkg{readr}) are ideal candidates.
\item Create a function in \pkg{fertile} with the same name that takes the same arguments (and always the dots \cmd{...}).
\item Write this new function so that it: 
\begin{enumerate}
\item Captures any arguments
\item Logs the name of the function called
\item Performs any checks on these arguments, and, finally:
\item Calls the original function with the original arguments. Except where warranted, the execution looks the same to the user as if they were calling the original function.
\end{enumerate}
\end{enumerate}
Most shims are quite simple and look something like what is shown below for \func{read\_csv}.

@@

In order to ensure that the \pkg{fertile} versions of functions ("shims") are always favored ("masked") over their original namesakes when called, \pkg{fertile} uses its own shims of the \func{library} and \func{require} functions to  manipulate the \R \func{search} path so that it is always located in the first position. In the \pkg{fertile} version of \func{library}, we detach \pkg{fertile} from the search path, load the requested package, and then re-attach \pkg{fertile}. This ensures that when a user executes a command, \R will check \pkg{fertile} for a matching function before considering other packages.

@@

<<is_file>>=
is_data_file(fs::path(proj_dir, "mice.csv"))
is_image_file(fs::path(proj_dir, "mice.csv"))
is_image_file(fs::path(proj_dir, "proteins_v_time.png"))
is_text_file(fs::path(proj_dir, "README.md"))
is_r_file(fs::path(proj_dir, "analysis.Rmd"))
@

@@

<<sandbox>>=
dir <- getwd()
new_dir <- sandbox(dir)

dir
fs::dir_ls(dir) %>% head(3)

new_dir
fs::dir_ls(new_dir) %>% head(3)
@

@@

<<proj_pkg_script, eval=TRUE, cache=TRUE>>=
install_script <- proj_pkg_script(proj_dir)
cat(readChar(install_script, 1e5))
@

@@

\pkg{fertile} is designed to: 
\begin{enumerate}
\item Be simple enough that users with minimal \R experience can use the package without issue.
\item Increase the reproducibility of work produced by its users, and:
\item Educate its users on why their work is or is not reproducible and provide guidance on how to address any problems.
\end{enumerate}

To test \pkg{fertile}'s effectiveness, we began an initial randomized control trial of the package on an introductory undergraduate data science course at Smith College in Spring 2020 \footnote{This study was approved by Smith College IRB, Protocol \#19-032}.

@@

Both groups are given a short quiz on different components of reproducibility that are intended to be taught by \pkg{fertile} at both the beginning and end of the semester. Their scores are then compared to see whether one group learned more than the other group or whether their scores were essentially equivalent. Additionally, for every homework assignment submitted, the professor takes note of whether or not the project compiles successfully.

@@

Based on the results, we could determine whether \pkg{fertile} was successful at achieving its intended goals:

\begin{enumerate}

\item A lack of notable difference between the \func{experimental} and \func{control} groups in terms of the number of code-related questions asked throughout the semester would indicate that \pkg{fertile} achieved its goal of simplicity.
\item A higher average for the \func{experimental} group in terms of the number of homework assignments that compiled successfully would indicate that \pkg{fertile} was successful in increasing reproducibility.
\item A greater increase over the semester in the reproducibility quiz scores for students in the \func{experimental} group compared with the \func{control} group would indicate that \pkg{fertile} achieved its goal of educating users on reproducibility.

\end{enumerate}

Success in achieving these goals would provide valuable evidence showing \pkg{fertile}'s benefit as tool to help educators introduce reproducibility concepts in the classrom.

 Unfortunately, we were unable to complete the analysis as intended as the trial had to be postponed after the COVID-19 pandemic significantly altered the experimental conditions at the midpoint of testing.
Although the experiment was unsuccessful in its first attempt, we hope to run the same trial again and gather data on \pkg{fertile}'s effectiveness.
